---
title: "Lincoln Data"
author: "Simon Woodward, DairyNZ 2018"
output: word_document
---

<!-- modified below cutting logic to use Katherine Tozer data -->

```{r setup, include=FALSE}
# include=FALSE (hide R code and output)
# echo=FALSE (hide R code)
# results=markup, results=asis, results=hide
# warning=FALSE, message=FALSE, error=TRUE
# fig.keep=none, fig.keep=all, fig.keep=high  
remove(list=ls())
knitr::opts_chunk$set(warning=FALSE, tidy=FALSE, fig.keep='all', fig.width=9)

# load libraries
# "lubridate" %in% (.packages()) # how to check if package loaded
library(tidyverse)
library(readxl)
library(stringr)
library(scales)
library(lubridate)
library(RColorBrewer)

# automatically convert text to something easy to work with
ensnakeify <- function(x) {
  x %>%
    iconv(to="ASCII//TRANSLIT") %>% # remove accents
    str_replace_na() %>% # convert NA to string
    str_to_lower() %>% # convert to lower case
    str_replace_all(pattern="[^[:alnum:]]", replacement=" ") %>% # convert non-alphanumeric to space
    str_trim() %>% # trim leading and trailing spaces
    str_replace_all(pattern="\\s+", replacement="_") # convert remaining spaces to underscore
}

# rounding function that works on dates
round_any = function(x, accuracy, f=round){
  if (is.Date(x)){
    switch(f,
           round=round_date(x, accuracy),
           ceiling=ceiling_date(x, accuracy),
           floor=floor_date(x, accuracy)
    )
  } else {
    f( x / accuracy) * accuracy
  }
}

# my_trans <- function(x){trans_new('my', 
#             transform = function(y){0.01*y},
#             inverse   = function(y){100*y}
#             )}

```

## Lincoln Data

```{r echo-FALSE}
# file name
file_name <- 'FD1004_C Data For Modelling.xlsx'

# seed rates in order
seed_rate_levels <- c('6kg', '12kg', '18kg', '24kg', '30kg')
```

## Rising Plate Meter

```{r include=FALSE}
# read and plot rpm data. There is some extra stuff in this sheet.
data_rpm <- read_xlsx(file_name, sheet='Canterbury RPM Heights')
names(data_rpm) <- ensnakeify(names(data_rpm))

# rename useful variables
data_rpm <- data_rpm %>%
  mutate(seed_rate = factor(seed_rate, levels=seed_rate_levels)) %>%
  rename(
       date_pre = date_pre_rpm,
       mass_pre = pregrazing_mass_kg_dm_ha,
       date_post = date_post_rpm,
       mass_post = postgrazing_mass_kg_dm_ha,
       date_grazed = date_grazed_d,
       grazing = graze_no
       ) %>%
  select(date_grazed, block, cultivar, seed_rate, grazing, date_pre, mass_pre, date_post, mass_post)

```

Average pre-graze mass = `r round(mean(data_rpm$mass_pre, na.rm=TRUE))`

Average post-graze mass = `r round(mean(data_rpm$mass_post, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot pre and post mass by cultivar, seed_rate and block
ybreaks <- seq(0, round_any(max(data_rpm$mass_pre, na.rm=TRUE), 2000, ceiling), 2000)
xbreaks <- seq(floor_date(min(data_rpm$date_pre), "years"), 
               ceiling_date(max(data_rpm$date_post), "years"), by="1 year")
data_rpm %>%
  split(.$block) %>%
  map(~ggplot(.) +
        labs(x='Date', y='Plate mass (kgDM/ha)', 
          title=paste('Pre and post plate mass (to ground level), Block', unique(.$block))) +
        theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
        geom_vline(mapping=aes(xintercept=date_grazed), colour='grey') +
        geom_point(mapping=aes(x=date_pre, y=mass_pre), colour='red') +
        geom_point(mapping=aes(x=date_post, y=mass_post), colour='blue') +
        facet_grid(cultivar ~ seed_rate) +
        scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
        scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
      )

```

<!-- ## Rising Plate Meter Growth Rate -->

<!-- ### Assumptions  -->

<!-- * Growth rate calculated from RPM data and assumed constant between grazings. -->

```{r include=FALSE, eval=FALSE}
# calculate growth rate before grazing by cultivar, seed_rate and block.
data_rpm <- data_rpm %>%
  group_by(block, cultivar, seed_rate) %>%
  arrange(date_grazed) %>%
  mutate(
    date_post_last = lag(date_post ,1), # FIXME could fail
    mass_post_last = lag(mass_post ,1), # FIXME could fail
    growth_days_pre = as.numeric(difftime(date_pre, date_post_last), units="days"),
    growth_pre = mass_pre - mass_post_last,
    growth_rate_pre = growth_pre / as.double(growth_days_pre),
    growth_rate_post = lead(growth_rate_pre, 1)
    ) 

```

<!-- Average growth rate = `r #round(mean(data_rpm$growth_rate_pre, na.rm=TRUE))` -->

```{r echo=FALSE, message=FALSE, results="hide", eval=FALSE}
# plot growth rate
ybreaks <- seq(round_any(min(data_rpm$growth_rate_pre, na.rm=TRUE), 25, floor), 
               round_any(max(data_rpm$growth_rate_pre, na.rm=TRUE), 25, ceiling), 25)
xbreaks <- seq(floor_date(min(data_rpm$date_pre), "years"), 
               ceiling_date(max(data_rpm$date_post), "years"), by="1 year")
data_rpm %>%
  split(.$block) %>%
  map(~ggplot(.) +
        labs(x='Date', y='Growth Rate (kgDM/ha/d)', 
             title=paste('Growth Rate (Previous), Block', unique(.$block))) +
        theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
        geom_vline(mapping=aes(xintercept=date_grazed), colour='grey') +
        geom_point(mapping=aes(x=date_grazed, y=growth_rate_pre), colour='magenta') +
        facet_grid(cultivar ~ seed_rate) +
        scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
        scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
      )

```

## Rising Plate Meter Harvest %

### Assumptions 

* Assume no pasture growth between pre and post RPM.

```{r include=FALSE}
# calculate harvest % by cultivar, seed_rate and block.
data_rpm <- data_rpm %>%
  group_by(block, cultivar, seed_rate) %>%
  mutate(
    delay_pre = as.numeric(difftime(date_grazed, date_pre), units="days"),
    delay_post = as.numeric(difftime(date_post, date_grazed), units="days"),
    mass_grazed = mass_pre, # + growth_rate_pre * delay_pre * 0,
    mass_resid = mass_post, # - growth_rate_post * delay_post * 0, 
    harv = (1 - mass_resid / mass_grazed) * 100 # proportion harvested
    )

```

Average harvest % = `r round(mean(data_rpm$harv, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot harvest %
ybreaks <- seq(0, round_any(max(data_rpm$harv, na.rm=TRUE), 20, ceiling), 20)
xbreaks <- seq(floor_date(min(data_rpm$date_pre), "years"), 
               ceiling_date(max(data_rpm$date_post), "years"), by="1 year")
data_rpm %>%
  split(.$block) %>%
  map(~ggplot(.) +
        labs(x='Date', y='Harvest %', 
             title=paste('Harvest %, Block', unique(.$block))) +
        theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
        geom_vline(mapping=aes(xintercept=date_grazed), colour='grey') +
        geom_point(mapping=aes(x=date_grazed, y=harv), colour='purple') +
        facet_grid(cultivar ~ seed_rate) +
        scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
        scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
      )

```

## Pasture Cut Mass

```{r include=FALSE}
# read cut mass data. Cuts were to 4-5 cm.
data_cut1 <- read_xlsx(file_name, sheet='Cut Yield Data Year1&2')
data_cut2 <- read_xlsx(file_name, sheet='Cut Yield Data Year3on Canty')
names(data_cut1) <- ensnakeify(names(data_cut1))
names(data_cut2) <- ensnakeify(names(data_cut2))

# rename useful variables
data_cut1 <- data_cut1 %>%
  mutate(seed_rate = factor(seed_rate, levels=seed_rate_levels)) %>%
  rename(
    date_cut = date_d,
    yield = yield_kg_dm_ha,
    grazing = graze_no
  ) %>%
  mutate(
    month = month(month)
  ) %>%
  group_by(date_cut, block, seed_rate, cultivar, grazing) %>%
  summarise(yield = mean(yield)) %>% 
  ungroup()

# rename useful variables
data_cut2 <- data_cut2 %>%
  mutate(seed_rate = factor(seed_rate, levels=seed_rate_levels)) %>%
  rename(
    date_cut = date_d,
    yield = yield_kg_dm_ha,
    dm_pc = dm_1, # dm_2 has a lot of missing
    grazing = graze_no
  ) %>%
  mutate(
    month = month(month)
  ) %>%
  select(date_cut, block, seed_rate, cultivar, yield, dm_pc, grazing)

# combine tables by row keeping common variables
data_cut <- full_join(data_cut1, data_cut2) 

# cutting dates for matching with other data sets
# date_cut <- unique(data_cut$date)
# closest_date_cut <- function(date){ # return index of closest date_cut
#   i <- which(abs(date-date_cut)==min(abs(date-date_cut)))[[1]]
# }

```

Average cut mass = `r round(mean(data_cut$yield, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot cut yield data
ybreaks <- seq(0, round_any(max(data_cut$yield, na.rm=TRUE), 1000, ceiling), 1000)
xbreaks <- seq(floor_date(min(data_cut$date_cut), "years"), 
               ceiling_date(max(data_cut$date_cut), "years"), by="1 year")
data_cut %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Cut mass (kgDM/ha)', 
         title=paste('Pasture Cut mass (pregrazing, cut to 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
    geom_point(mapping=aes(x=date_cut, y=yield), colour='red') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
  )

```

## Estimate Mass Below Cutting Height

### Assumptions 

* Total mass at cutting assumed to be equal to pregraze RPM.

```{r include=FALSE}
# find cut matching rpm data
# data_rpm$date_cut_i <- lapply(data_rpm$date_pre, closest_date_cut)
# data_rpm$date_cut <- date_cut[unlist(data_rpm$date_cut_i)]

# join cut and rpm data by graing and estimate mass on cutting date
# mass below cutting height is this minus cut yield
# data_cut$date_cut <- data_cut$date
data_bc <- data_rpm %>%
  left_join(data_cut, by=c('block', 'seed_rate', 'cultivar', 'grazing')) %>%
  select(date_pre, date_cut, block, seed_rate, cultivar, mass_pre, yield, grazing) %>%
  mutate(delay = as.integer(as.Date(date_cut) - as.Date(date_pre))) %>% 
  # filter(abs(delay)<4) %>%
  # drop_na() %>%
  mutate(
    mass_cut = mass_pre, # + growth_rate_pre * delay * 0,
    below = mass_cut - yield
    )
unique(data_bc$delay) # check
   
# copy below back into data_cut
temp <- data_cut %>%
  left_join(data_bc, by=c('block', 'seed_rate', 'cultivar', 'grazing')) %>%
  select(date_cut.x, block, seed_rate, cultivar, yield.x, dm_pc, grazing,
         date_pre, mass_pre, delay, below) %>%
  rename(date_cut = date_cut.x,
         yield = yield.x)
  
data_cut <- temp

```

Average mass below cutting estimate = `r round(mean(data_cut$below, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot below cutting mass
ybreaks <- seq(0, round_any(max(data_bc$below, na.rm=TRUE), 1000, ceiling), 1000)
xbreaks <- seq(floor_date(min(data_bc$date_cut, na.rm=TRUE), "years"), 
               ceiling_date(max(data_bc$date_cut, na.rm=TRUE), "years"), by="1 year")
data_bc %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight (kgDM/ha)', fill='Species',
         title=paste('Mass Below Cut (below 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_point(mapping=aes(x=date_cut, y=below), colour='brown') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
  )

```

## Pasture Cuts DM%

Average cut dry matter % = `r round(mean(data_cut$dm_pc, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot DM% data
ybreaks <- seq(0, round_any(max(data_cut$dm_pc, na.rm=TRUE), 10, ceiling), 10)
xbreaks <- seq(floor_date(min(data_cut$date_cut, na.rm=TRUE), "years"), 
               ceiling_date(max(data_cut$date_cut, na.rm=TRUE), "years"), by="1 year")
data_cut %>%
  split(.$block) %>%
  map(~ggplot(.) +
  labs(x='Date', y='Dry matter %', 
       title=paste('Pregazing Dry Matter % (cut to 4cm), Block', unique(.$block))) +
  theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
  geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
  geom_point(mapping=aes(x=date_cut, y=dm_pc), colour='purple') +
  facet_grid(cultivar ~ seed_rate ) +
  scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
  scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
  )

```

## Tiller Density

```{r include=FALSE}
# read tiller data
data_till <- read_xlsx(file_name, sheet='Tiller density Canterbury')
names(data_till) <- ensnakeify(names(data_till))

# rename useful variables
data_till <- data_till %>%
  mutate(seed_rate = factor(seed_rate, levels=seed_rate_levels)) %>%
  rename(
    date_till = date_d,
    tillers = ryegrass_tiller_density_tillers_m2
  ) %>%
  group_by(block, cultivar, seed_rate, date_till) %>%
  summarise(mean_tillers = mean(tillers)) %>% 
  ungroup()

```

Average tiller density = `r round(mean(data_till$mean_tillers, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot tiller data 
ybreaks <- seq(0, round_any(max(data_till$mean_tillers, na.rm=TRUE), 2500, ceiling), 2500)
xbreaks <- seq(floor_date(min(data_till$date_till), "years"), 
               ceiling_date(max(data_till$date_till), "years"), by="1 year")
data_till %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Tiller density'~(tillers~m^{2}), 
         title=paste('Tiller Density, Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
    geom_point(mapping=aes(x=date_till, y=mean_tillers), colour='darkgreen') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, 
                       limits=c(ybreaks[1], round_any(max(data_till$mean_tillers, na.rm=TRUE), 1000, ceiling)))
  )

```

## Botanical Composition

```{r echo=FALSE, message=FALSE, results="hide"}
# read botanical data. Skip blank line at top
data_bot <- read_xlsx(file_name, sheet='Botanical Composition Data', skip=1)
names(data_bot) <- ensnakeify(names(data_bot))

# rename useful variables
data_bot <- data_bot %>%
  mutate(seed_rate = factor(seed_rate, levels=seed_rate_levels)) %>%
  rename(
    date_bot = date_d,
    leaf = perennial_ryegrass_leaf, # (includes pseudostem)
    stem = perennial_ryegrass_reproductive_stem,
    ann = annual_ryegrass,
    wc = white_clover,
    poa = poa_sp,
    ogx = other_grasses_excluding_poa,
    og = other_grasses_including_poa,
    weed = weeds,
    dead = dead
    ) %>%
  mutate(
    total = leaf + stem + ann + wc + og + weed + dead, # should always be 100
    rgfrac = 100 * (leaf + stem) / (leaf + stem + ann + wc + og + weed), # of green
    month = month(dmy(paste('01', month, '2011'))),
    # estimate botanicals composition below cutting height
    x = month/12,
    y = 129.07*x^6 - 416.62*x^5 + 510.83*x^4 - 291.95*x^3 + 75.37*x^2 - 6.95*x + 1.32,
    leaf_below = leaf / y, # estimated from  Tozer data (includes pseudostem)
    stem_below = stem / 5 # estimated from  Tozer data
  ) %>%
  select(leaf, stem, leaf_below, stem_below, ann, wc, og, weed, dead, date_bot, total, block, seed_rate, cultivar, rgfrac) 
  
# add grazing number
temp_cut <- data_cut %>% 
  distinct(block, cultivar, seed_rate, date_cut) 
closest_cut <- function(block_bot, cultivar_bot, seed_rate_bot, date_bot){
  temp <- temp_cut %>% 
    filter(block==block_bot, cultivar==cultivar_bot, seed_rate==seed_rate_bot)
  ans <- date_bot
  for (i in 1:length(date_bot)){
    delay <- date_bot[i] - temp$date_cut
    ans[i] <- temp$date_cut[which(abs(delay)==min(abs(delay)))[[1]]]
  }
  return(ans)
}
temp_bot <- data_bot %>% 
  group_by(block, cultivar, seed_rate, date_bot) %>%
  mutate(date_cut = closest_cut(block, cultivar, seed_rate, date_bot)) %>% 
  ungroup %>% 
  left_join(select(data_cut, block, cultivar, seed_rate, date_cut, grazing),
            by = c("block", "cultivar", "seed_rate", "date_cut")) %>% 
  mutate(delay = as.integer(as.Date(date_bot) - as.Date(date_cut)),
         date_cut = if_else(abs(delay) < 7, date_cut, as.POSIXct(NA)),
         grazing = if_else(abs(delay) < 7, grazing, NA_real_))
  
data_bot <- temp_bot

# gather
data_bot2 <- data_bot %>%
  gather(leaf, stem, ann, wc, og, weed, dead, key='species', value='fraction') %>%
  mutate(species = factor(species, 
                          levels=c('leaf', 'stem', 'ann', 'og', 'wc', 'weed', 'dead'))) 

# for debugging
write_tsv(data_bot2, "data_bot2_Lincoln.tsv")

# plot botanical data
ybreaks <- seq(0, 100, 20)
xbreaks <- seq(floor_date(min(data_bot2$date_bot), "years"), 
               ceiling_date(max(data_bot2$date_bot), "years"), by="1 year")
data_bot2 %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight %', fill='Species',
         title=paste('Botanical composition (cut to 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_bar(mapping=aes(x=date_bot, y=fraction, fill=species), stat='identity') +
    scale_fill_brewer(palette='Paired') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=range(xbreaks)) +
    scale_y_continuous(breaks=ybreaks, limits=c(0, 120))
  )

```

## Ryegrass Fraction

### Assumptions

* Ryegrass fraction calculated on green mass only.

```{r echo=FALSE, message=FALSE, results="hide"}
# plot ryegrass fraction
ybreaks <- seq(0, 100, 20)
xbreaks <- seq(floor_date(min(data_bot$date_bot), "years"), 
               ceiling_date(max(data_bot$date_bot), "years"), by="1 year")
data_bot %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight %', 
         title=paste('Ryegrass Fraction (above 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_point(mapping=aes(x=date_bot, y=rgfrac), colour="darkgreen") +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=range(xbreaks)) +
    scale_y_continuous(breaks=ybreaks, limits=c(0, 120))
  )

```

## Estimate Botancial Mass

### Assumptions 

* Total mass at botanical date assumed to be equal to pregraze RPM.

* Cut mass at botanical date assumed to be equal to cut yield.

* Botanical composition below cutting height estimated from Tozer data.

```{r echo=FALSE, message=FALSE, results="hide"}
# it's not clear how best to extrapolate botanical % above 4 cm to botancial mass above ground
# find cut matching botanical data
# data_bot$date_cut_i <- lapply(data_bot$date, closest_date_cut)
# data_bot$date_cut <- date_cut[unlist(data_bot$date_cut_i)]

# join cut and botanical data by date_cut
# data_cut$date_cut <- data_cut$date
data_bm <- data_bot %>%
  # rename(date_bot = date) %>%
  left_join(data_cut, by=c('grazing', 'date_cut', 'block', 'seed_rate', 'cultivar')) %>%
  select(date_bot, block, seed_rate, cultivar, 
         leaf, stem, ann, wc, og, weed, dead, leaf_below, stem_below,
         date_cut, grazing, yield, below) %>%
  rename(yield_cut = yield) %>%
  mutate(
    yield_bot = yield_cut #+ below * 0 + growth_rate_pre * delay * 0 # assumed on date_bot, not including below mass
    )
         
# gather
data_bm2 <- data_bm %>%
  gather(leaf, stem, ann, wc, og, weed, dead, key='species', value='fraction') %>%
  mutate(
    species = factor(species, levels=c('leaf', 'stem', 'ann', 'og', 'wc', 'weed', 'dead')),
    species_mass = fraction / 100 * yield_bot # include mass below cutting?
    ) 

# plot botanical data
ybreaks <- seq(0, round_any(max(data_bm2$yield_bot, na.rm=TRUE), 1000, ceiling), 1000)
ybreaks2 <- seq(0, round_any(max(data_bm2$yield_bot, na.rm=TRUE), 100, ceiling), 100)
xbreaks <- seq(floor_date(min(data_bm2$date_bot), "years"), 
               ceiling_date(max(data_bm2$date_bot), "years"), by="1 year")
data_bm2 %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight (kgDM/ha)', fill='Species',
         title=paste('Herbage Mass Fractions (above 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_point(mapping=aes(x=date_bot, y=yield_bot), colour='grey', size=2, shape=1) +
    geom_bar(mapping=aes(x=date_bot, y=species_mass, fill=species), stat='identity') +
    scale_fill_brewer(palette='Paired') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=range(xbreaks)) +
    scale_y_continuous(breaks=ybreaks, limits=range(ybreaks2))
  )

```

<!-- ## Light Interception -->

```{r include=FALSE, eval=FALSE}
# read li data
data_li <- read_xlsx(file_name, sheet='Waikato LightInterception')
names(data_li) <- ensnakeify(names(data_li))

# rename useful variables
data_li <- data_li %>%
  mutate(
    block = floor((plot-1)/20)+1, # this needs to be checked 
    seed_rate = factor(seed_rate, levels=seed_rate_levels)
    ) %>%
  rename(
    date_li = date,
    date_grazed = date_last_grazed,
    li = light_interception
  ) %>%
  select(date_li, block, seed_rate, cultivar, li, date_grazed)

```

<!-- Average light interception = r # round(mean(data_li$li, na.rm=TRUE)) -->

```{r echo=FALSE, message=FALSE, results="hide", eval=FALSE}
# plot light interception data
ybreaks <- seq(0, round_any(max(data_li$li, na.rm=TRUE), 25, ceiling), 25)
xbreaks <- seq(floor_date(min(data_li$date_li), "months"), 
               ceiling_date(max(data_li$date_li), "months"), by="1 month")
data_li %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Light interception %', 
         title=paste('Light Interception 2011, Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
    geom_point(mapping=aes(x=date_li, y=li), colour='magenta') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=format(xbreaks, "%b-%Y"), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))

  )

```

## Soil Moisture

```{r include=FALSE}
# read soil moisture data
data_sm <- read_xlsx(file_name, sheet='SoilMoist', skip=9, 
                     col_names = c("date_sm", "rn", 
                    "sm5cm_av", "sm10cm_av", 
                    "sm5cm_max", "sm10cm_max", 
                    "sm5cm_min", "sm10cm_min")) 

# rename useful variables
data_sm <- data_sm %>%
  # mutate(seed_rate = factor(seed_rate, levels=seed_rate_levels)) %>%
  mutate(
    # date = date_measured_d,
    # sm = soil_moisture
    seed_rate = factor("18kg", levels=seed_rate_levels),
    block = 3,
    cultivar = "Alto",
    dom = day(date_sm),
    sm = sm10cm_av * 100 ,
    mean_sm = sm10cm_av * 100
  ) %>% 
  filter(dom==1) # only take one point per month
  # ) %>%
  # group_by(block, cultivar, seed_rate, date) %>%
  # summarise(mean_sm = mean(sm))
data_sm <- rbind(mutate(data_sm, block=1), 
                 mutate(data_sm, block=2), 
                 mutate(data_sm, block=3), 
                 mutate(data_sm, block=4), 
                 mutate(data_sm, block=5)) 

```

Average soil moisture = `r round(mean(data_sm$mean_sm, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot sol moisture data
ybreaks <- seq(0, round_any(max(data_sm$mean_sm, na.rm=TRUE), 20, ceiling), 20)
xbreaks <- seq(floor_date(min(data_sm$date_sm), "years"), 
               ceiling_date(max(data_sm$date_sm), "years"), by="1 year")

ggplot(filter(data_sm, block==3, seed_rate=="18kg", cultivar=="Alto")) +
    labs(x='Date', y='Soil Moisture %', 
         title=paste('Soil Moisture, All Blocks')) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_vline(data=filter(data_rpm, block==3, seed_rate=="18kg", cultivar=="Alto"),
               mapping=aes(xintercept=date_grazed), colour='grey') +
    geom_point(mapping=aes(x=date_sm, y=mean_sm), colour='blue') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=format(xbreaks, "%b-%Y"), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))

```

<!-- ## Calculations -->

```{r include=FALSE}
# additional calculations to prepare data for model calibration
# move post harvest measurements if on grazing day since model does harvest at start of day
data_rpm$date_post2 <- as.Date(with(data_rpm,
                            ifelse(date_post > date_grazed, 
                                   as.Date(date_post), 
                                   as.Date(date_post) + 1)
                            ), origin="1970-01-01")

# some calculations
data_rpm <- data_rpm %>%
  mutate(
    year_pre = year(date_pre),
    doy_pre = yday(date_pre),
    year_post = year(date_post2),
    doy_post = yday(date_post2),
    year_grazed = year(date_grazed),
    doy_grazed = yday(date_grazed), 
    days_pre = difftime(date_grazed, date_pre),
    days_post = difftime(date_post2, date_grazed)
    )

```

## Write Harvest Dates and Data for Model Calibration

```{r include=FALSE, result="hide"}
# choose data
acultivar <- 'Alto' # only Alto and Halo have light interception data
aseed_rate <- '18kg'
calib_start <- ymd("20120401") # period for data weight = 1
calib_end <- ymd("20171231") # period for data weight = 1
ablock <- 3

for (ablock in c(1,2,3,4,5)) { # loop through data sets

# write harvest dates and harvest % for selected series
file_name <- paste("harvest_Lincoln_", ablock, ".txt", sep="")
data_h <- data_rpm %>%
  select(block, seed_rate, cultivar, year_grazed, doy_grazed, harv) %>%
  filter(block == ablock & seed_rate==aseed_rate & cultivar==acultivar) %>%
  drop_na()
days_harvest <- matrix(as.integer(-1), nrow=100, ncol=3) # up to 100 harvests
days_harvest[1:nrow(data_h),] <- c(data_h$year_grazed, data_h$doy_grazed, data_h$harv)
write.table(days_harvest, file=file_name, row.names=FALSE, col.names=FALSE, sep='\t')

# write calibration data
file_name <- paste("data_calibration_Lincoln_", ablock, ".txt", sep="")

# collect the data in this list
data_c <- vector("list", 7) 
err_c <- c(TILTOTI=400, CLV=10, CST=4, WCLM=2, BASAL=2)
err_c <- c(TILTOTI=1000, CLV=40, CST=5, WCLM=5, BASAL=10)

# pre and post mass (but this includes other species!)
# temp <- data_rpm %>%
#   select(block, seed_rate, cultivar, mass_pre, year_pre, doy_pre, 
#          mass_post, year_post, doy_post) %>%
#   filter(block==ablock & seed_rate==aseed_rate & cultivar==acultivar) 
# data_c[[1]] <- with(temp, tibble(var='DM', year=year_pre, 
#                                  doy=doy_pre, data=mass_pre/10) %>% drop_na())
# data_c[[2]] <- with(temp, tibble(var='DM', year=year_post, 
#                                  doy=doy_post, data=mass_post/10) %>% drop_na())

# ryegrass tillers
temp <- data_till %>%
  select(block, seed_rate, cultivar, mean_tillers, date_till) %>%
  filter(block==ablock & seed_rate==aseed_rate & cultivar==acultivar) 
data_c[[3]] <- with(temp, tibble(var='TILTOT', year=year(date_till), 
                                 doy=yday(date_till), data=mean_tillers, sd=err_c['TILTOTI'], type="sd",
                                 weight=ifelse(((date_till>=calib_start)&(date_till<=calib_end)), 1, 0)) %>% drop_na())

# ryegrass mass (total or above cutting height? depending on definition of yield_bot)
temp <- data_bm %>%
  # rename(date = date_cut) %>%
  select(block, seed_rate, cultivar, leaf, stem, leaf_below, stem_below, yield_bot, below, date_bot) %>%
  filter(block==ablock & seed_rate==aseed_rate & cultivar==acultivar) 
data_c[[4]] <- with(temp, tibble(var='CLV', year=year(date_bot), 
                                 doy=yday(date_bot), data=leaf/100*yield_bot/10*0.45+leaf_below/100*below/10*0.45, 
                                 sd=err_c['CLV'], type="sd",
                                 weight=ifelse(((date_bot>=calib_start)&(date_bot<=calib_end)), 1, 0)) %>% drop_na())
data_c[[5]] <- with(temp, tibble(var='CST', year=year(date_bot), 
                                 doy=yday(date_bot), data=stem/100*yield_bot/10*0.45+stem_below/100*below/10*0.45, 
                                 sd=err_c['CST'], type="sd",
                                 weight=ifelse(((date_bot>=calib_start)&(date_bot<=calib_end)), 1, 0)) %>% drop_na())

# light interception (but this includes all species!)
# temp <- data_li %>%
#   select(block, seed_rate, cultivar, li, date) %>%
#   filter(block==ablock & seed_rate==aseed_rate & cultivar==acultivar) 
# data_c[[5]] <- with(temp, tibble(var='LINT', year=year(date), 
#                                  doy=yday(date), data=li) %>% drop_na())

# soil moisture
temp <- data_sm %>%
  select(block, seed_rate, cultivar, mean_sm, date_sm) %>%
  filter(block==ablock & seed_rate==aseed_rate & cultivar==acultivar)
data_c[[6]] <- with(temp, tibble(var='WCLM', year=year(date_sm),
                                 doy=yday(date_sm), data=mean_sm, sd=err_c['WCLM'], type="sd",
                                 weight=ifelse(((date_sm>=calib_start)&(date_sm<=calib_end)), 1, 0)) %>% drop_na())

# ryegrass fraction of green
# temp <- data_bot %>%
#   select(block, seed_rate, cultivar, rgfrac, date_bot) %>%
#   filter(block==ablock & seed_rate==aseed_rate & cultivar==acultivar) 
# data_c[[7]] <- with(temp, tibble(var='BASAL', year=year(date_bot), 
#                                  doy=yday(date_bot), data=rgfrac, sd=err_c['BASAL'], type="sd",
#                                  weight=ifelse(((date_bot>=calib_start)&(date_bot<=calib_end)), 1, 0)) %>% drop_na())

# bind list and write file
data_calib <- bind_rows(data_c)
data_calib <- arrange(data_calib, var, year, doy)
write.table(data_calib, file=file_name, row.names=FALSE, col.names=FALSE, sep='\t', quote=FALSE)

} # end data set loop

```

