---
output: word_document
---

<!-- modified below cutting logic to use Katherine Tozer data -->

```{r setup, include=FALSE}
# include=FALSE (hide R code and output)
# echo=FALSE (hide R code)
# results=markup, results=asis, results=hide
# warning=FALSE, message=FALSE, error=TRUE
# fig.keep=none, fig.keep=all, fig.keep=high  
knitr::opts_chunk$set(warning=FALSE, tidy=FALSE, fig.keep='all', fig.width=9)

# load libraries
library(scales)
library(RColorBrewer)
library(tidyverse)
library(readxl)
library(stringr)
library(lubridate)

# get data
# load(paste(site, ".RData", sep=""))

```

---
title: "`r paste(site, "Data")`"
author: "`r paste("Simon Woodward, DairyNZ,", time_stamp)`"
---

## Rising Plate Meter

Average pre-graze mass = `r round(mean(data_rpm$mass_pre, na.rm=TRUE))`

Average post-graze mass = `r round(mean(data_rpm$mass_post, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot pre and post mass by cultivar, seed_rate and block
ybreaks <- seq(0, round_any(max(data_rpm$mass_pre, na.rm=TRUE), 2000, ceiling), 2000)
xbreaks <- seq(floor_date(min(data_rpm$date_pre, na.rm=TRUE), "years"), 
               ceiling_date(max(data_rpm$date_post, na.rm=TRUE), "years"), by="1 year")
data_rpm %>%
  split(.$block) %>%
  map(~ggplot(.) +
        labs(x='Date', y='Plate mass (kgDM/ha)', 
          title=paste('Pre and post plate mass (to ground level), Block', unique(.$block))) +
        theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
        geom_vline(mapping=aes(xintercept=date_grazed), colour='grey') +
        geom_point(mapping=aes(x=date_pre, y=mass_pre), colour='red') +
        geom_point(mapping=aes(x=date_post, y=mass_post), colour='blue') +
        facet_grid(cultivar ~ seed_rate) +
        scale_x_date(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
        scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
      )
```

<!-- ## Rising Plate Meter Growth Rate -->

<!-- Growth rate calculated from RPM data and assumed constant between grazings -->

<!-- Average growth rate = `r # round(mean(data_rpm$growth_rate_pre, na.rm=TRUE))` -->

```{r echo=FALSE, message=FALSE, results="hide", eval=FALSE}
# plot growth rate
ybreaks <- seq(round_any(min(data_rpm$growth_rate_pre, na.rm=TRUE), 25, floor), 
               round_any(max(data_rpm$growth_rate_pre, na.rm=TRUE), 25, ceiling), 25)
xbreaks <- seq(floor_date(min(data_rpm$date_pre, na.rm=TRUE), "years"), 
               ceiling_date(max(data_rpm$date_post, na.rm=TRUE), "years"), by="1 year")
data_rpm %>%
  split(.$block) %>%
  map(~ggplot(.) +
        labs(x='Date', y='Growth Rate (kgDM/ha/d)', 
             title=paste('Growth Rate (Previous), Block', unique(.$block))) +
        theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
        geom_vline(mapping=aes(xintercept=date_grazed), colour='grey') +
        geom_point(mapping=aes(x=date_grazed, y=growth_rate_pre), colour='magenta') +
        facet_grid(cultivar ~ seed_rate) +
        scale_x_date(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
        scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
      )
```

## Pasture Cut Mass

Average cut mass = `r round(mean(data_cut$yield_cut, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot cut yield data
ybreaks <- seq(0, round_any(max(data_cut$yield_cut, na.rm=TRUE), 1000, ceiling), 1000)
xbreaks <- seq(floor_date(min(data_cut$date_cut, na.rm=TRUE), "years"), 
               ceiling_date(max(data_cut$date_cut, na.rm=TRUE), "years"), by="1 year")
data_cut %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Cut mass (kgDM/ha)', 
         title=paste('Pasture Cut mass (pregrazing, cut to 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
    geom_point(mapping=aes(x=date_cut, y=yield_cut), colour='red') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_date(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
  )
```

## Pasture Cuts DM%

Average cut dry matter % = `r round(mean(data_cut$dm_cut, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide", eval=FALSE}
# plot DM% data
ybreaks <- seq(0, round_any(max(data_cut$dm_cut, na.rm=TRUE), 10, ceiling), 10)
xbreaks <- seq(floor_date(min(data_cut$date_cut, na.rm=TRUE), "years"), 
               ceiling_date(max(data_cut$date_cut, na.rm=TRUE), "years"), by="1 year")
data_cut %>%
  split(.$block) %>%
  map(~ggplot(.) +
  labs(x='Date', y='Dry matter %', 
       title=paste('Pregazing Dry Matter % (cut to 4cm), Block', unique(.$block))) +
  theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
  geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
  geom_point(mapping=aes(x=date_cut, y=dm_cut), colour='purple') +
  facet_grid(cultivar ~ seed_rate ) +
  scale_x_date(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
  scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
  )
```

## Tiller Density

Average tiller density = `r round(mean(data_till$tillers, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot tiller data 
ybreaks <- seq(0, round_any(max(data_till$tillers, na.rm=TRUE), 5000, ceiling), 5000)
xbreaks <- seq(floor_date(min(data_till$date_till, na.rm=TRUE), "years"), 
               ceiling_date(max(data_till$date_till, na.rm=TRUE), "years"), by="1 year")
data_till %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Tiller density'~(tillers~m^{2}), 
         title=paste('Tiller Density, Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
    geom_point(mapping=aes(x=date_till, y=tillers), colour='darkgreen') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_date(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, 
                       limits=c(ybreaks[1], round_any(max(data_till$tillers, na.rm=TRUE), 1000, ceiling)))
  )
```

## Botanical Composition

```{r echo=FALSE, message=FALSE, results="hide"}
# plot botanical data
data_bot2 <- data_bot2 %>% 
  mutate(date_bot = make_date(year(date_bot), month(date_bot), 1))
ybreaks <- seq(0, 100, 20)
xbreaks <- seq(floor_date(min(data_bot2$date_bot, na.rm=TRUE), "years"), 
               ceiling_date(max(data_bot2$date_bot, na.rm=TRUE), "years"), by="1 year")
data_bot2 %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight %', fill='Species',
         title=paste('Botanical composition (cut to 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_bar(mapping=aes(x=date_bot, y=fraction, fill=species), stat='identity') +
    scale_fill_brewer(palette='Paired') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_date(breaks=xbreaks, labels=year(xbreaks), limits=range(xbreaks)) + # FIXME drops data?
    scale_y_continuous(breaks=ybreaks, limits=c(0, 120))
  )
```

## Ryegrass Fraction 

Ryegrass fraction calculated on green mass only

```{r echo=FALSE, message=FALSE, results="hide"}
# plot ryegrass fraction
ybreaks <- seq(0, 100, 20)
xbreaks <- seq(floor_date(min(data_bot$date_bot, na.rm=TRUE), "years"), 
               ceiling_date(max(data_bot$date_bot, na.rm=TRUE), "years"), by="1 year")
data_bot %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight %', 
         title=paste('Ryegrass Fraction (above 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_point(mapping=aes(x=date_bot, y=rgfrac), colour="darkgreen") +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_date(breaks=xbreaks, labels=year(xbreaks), limits=range(xbreaks)) +
    scale_y_continuous(breaks=ybreaks, limits=c(0, 120))
  )
```

## Soil Moisture

Average soil moisture = `r round(mean(data_sm$mean_sm, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot soil moisture data
ybreaks <- seq(0, round_any(max(data_sm$mean_sm, na.rm=TRUE), 20, ceiling), 20)
xbreaks <- seq(floor_date(min(data_sm$date_sm, na.rm=TRUE), "years"), 
               ceiling_date(max(data_sm$date_sm, na.rm=TRUE), "years"), by="1 year")
data_sm %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Soil Moisture %', 
         title=paste('Soil Moisture 2017, Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
    geom_point(mapping=aes(x=date_sm, y=mean_sm), colour='blue') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_date(breaks=xbreaks, labels=format(xbreaks, "%b-%Y"), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
  )
```

## Mass Below Cutting Height

Estimated using linear regression of Cut Mass on RPM 

Average mass below cutting estimate = `r round(mean(data_bm$yield_below, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot below cutting mass
ybreaks <- seq(0, round_any(max(data_bm$yield_below, na.rm=TRUE), 1000, ceiling), 1000)
xbreaks <- seq(floor_date(min(data_bm$date_cut, na.rm=TRUE), "years"), 
               ceiling_date(max(data_bm$date_cut, na.rm=TRUE), "years"), by="1 year")
data_bm %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight (kgDM/ha)', fill='Species',
         title=paste('Mass Below Cut (below 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_point(mapping=aes(x=date_cut, y=yield_below), colour='brown') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_date(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
  )
```

## Harvest %

Estimated using linear regression of Tozer data on RPM

Average harvest % = `r round(mean(data_rpm$harv, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot harvest %
ybreaks <- seq(0, round_any(max(data_rpm$harv, na.rm=TRUE), 20, ceiling), 20)
xbreaks <- seq(floor_date(min(data_rpm$date_pre, na.rm=TRUE), "years"), 
               ceiling_date(max(data_rpm$date_post, na.rm=TRUE), "years"), by="1 year")
data_rpm %>%
  split(.$block) %>%
  map(~ggplot(.) +
        labs(x='Date', y='Harvest %', 
             title=paste('Harvest %, Block', unique(.$block))) +
        theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
        geom_vline(mapping=aes(xintercept=date_grazed), colour='grey') +
        geom_point(mapping=aes(x=date_grazed, y=harv), colour='purple') +
        facet_grid(cultivar ~ seed_rate) +
        scale_x_date(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
        scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
      )
```

## Botanical Mass

Estimated using Tozer data

```{r echo=FALSE, message=FALSE, results="hide"}
# plot botanical data
data_bm2 <- data_bm2 %>% 
  mutate(date_bot = make_date(year(date_bot), month(date_bot), 1))
ybreaks <- seq(0, round_any(max(data_bm2$yield_below, na.rm=TRUE), 1000, ceiling), 1000)
ybreaks2 <- seq(0, round_any(max(data_bm2$yield_below, na.rm=TRUE), 100, ceiling), 100)
xbreaks <- seq(floor_date(min(data_bm2$date_bot, na.rm=TRUE), "years"), 
               ceiling_date(max(data_bm2$date_bot, na.rm=TRUE), "years"), by="1 year")
data_bm2 %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight (kgDM/ha)', fill='Species',
         title=paste('Herbage Mass Fractions, Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_point(mapping=aes(x=date_bot, y=yield_above), colour='grey', size=2, shape=1) +
    geom_bar(mapping=aes(x=date_bot, y=species_mass, fill=species), stat='identity') +
    scale_fill_brewer(palette='Paired') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_date(breaks=xbreaks, labels=year(xbreaks), limits=range(xbreaks)) +
    scale_y_continuous(breaks=ybreaks, limits=range(ybreaks2))
  )
```

