---
title: "Northland Data"
author: "Simon Woodward, DairyNZ 2018"
output: word_document
---

<!-- modified below cutting logic to use Katherine Tozer data -->

```{r setup, include=FALSE}
# include=FALSE (hide R code and output)
# echo=FALSE (hide R code)
# results=markup, results=asis, results=hide
# warning=FALSE, message=FALSE, error=TRUE
# fig.keep=none, fig.keep=all, fig.keep=high  
knitr::opts_chunk$set(warning=FALSE, tidy=FALSE, fig.keep='all', fig.width=9)

# load libraries
library(scales)
library(RColorBrewer)
library(tidyverse)
library(readxl)
library(stringr)
library(lubridate)

# get data
load("Northland.RData")

```

## Rising Plate Meter

Average pre-graze mass = `r round(mean(data_rpm$mass_pre, na.rm=TRUE))`

Average post-graze mass = `r round(mean(data_rpm$mass_post, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot pre and post mass by cultivar, seed_rate and block
ybreaks <- seq(0, round_any(max(data_rpm$mass_pre, na.rm=TRUE), 2000, ceiling), 2000)
xbreaks <- seq(floor_date(min(data_rpm$date_pre, na.rm=TRUE), "years"), 
               ceiling_date(max(data_rpm$date_post), "years"), by="1 year")
data_rpm %>%
  split(.$block) %>%
  map(~ggplot(.) +
        labs(x='Date', y='Plate mass (kgDM/ha)', 
          title=paste('Pre and post plate mass (to ground level), Block', unique(.$block))) +
        theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
        geom_vline(mapping=aes(xintercept=date_grazed), colour='grey') +
        geom_point(mapping=aes(x=date_pre, y=mass_pre), colour='red') +
        geom_point(mapping=aes(x=date_post, y=mass_post), colour='blue') +
        facet_grid(cultivar ~ seed_rate) +
        scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
        scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
      )
```

<!-- ## Rising Plate Meter Growth Rate -->

<!-- Growth rate calculated from RPM data and assumed constant between grazings -->

<!-- Average growth rate = `r # round(mean(data_rpm$growth_rate_pre, na.rm=TRUE))` -->

```{r echo=FALSE, message=FALSE, results="hide", eval=FALSE}
# plot growth rate
ybreaks <- seq(round_any(min(data_rpm$growth_rate_pre, na.rm=TRUE), 25, floor), 
               round_any(max(data_rpm$growth_rate_pre, na.rm=TRUE), 25, ceiling), 25)
xbreaks <- seq(floor_date(min(data_rpm$date_pre, na.rm=TRUE), "years"), 
               ceiling_date(max(data_rpm$date_post), "years"), by="1 year")
data_rpm %>%
  split(.$block) %>%
  map(~ggplot(.) +
        labs(x='Date', y='Growth Rate (kgDM/ha/d)', 
             title=paste('Growth Rate (Previous), Block', unique(.$block))) +
        theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
        geom_vline(mapping=aes(xintercept=date_grazed), colour='grey') +
        geom_point(mapping=aes(x=date_grazed, y=growth_rate_pre), colour='magenta') +
        facet_grid(cultivar ~ seed_rate) +
        scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
        scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
      )
```

## Rising Plate Meter Harvest %

Assume no pasture growth between pre and post RPM

Average harvest % = `r round(mean(data_rpm$harv, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot harvest %
ybreaks <- seq(0, round_any(max(data_rpm$harv, na.rm=TRUE), 20, ceiling), 20)
xbreaks <- seq(floor_date(min(data_rpm$date_pre, na.rm=TRUE), "years"), 
               ceiling_date(max(data_rpm$date_post, na.rm=TRUE), "years"), by="1 year")
data_rpm %>%
  split(.$block) %>%
  map(~ggplot(.) +
        labs(x='Date', y='Harvest %', 
             title=paste('Harvest %, Block', unique(.$block))) +
        theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
        geom_vline(mapping=aes(xintercept=date_grazed), colour='grey') +
        geom_point(mapping=aes(x=date_grazed, y=harv), colour='purple') +
        facet_grid(cultivar ~ seed_rate) +
        scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
        scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
      )
```

## Pasture Cut Mass

Average cut mass = `r round(mean(data_cut$yield, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot cut yield data
ybreaks <- seq(0, round_any(max(data_cut$yield, na.rm=TRUE), 1000, ceiling), 1000)
xbreaks <- seq(floor_date(min(data_cut$date_cut), "years"), 
               ceiling_date(max(data_cut$date_cut), "years"), by="1 year")
data_cut %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Cut mass (kgDM/ha)', 
         title=paste('Pasture Cut mass (pregrazing, cut to 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
    geom_point(mapping=aes(x=date_cut, y=yield), colour='red') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
  )
```

## Estimate Mass Below Cutting Height

Total mass at cutting assumed to be equal to pregraze RPM

Average mass below cutting estimate = `r round(mean(data_cut$below, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot below cutting mass
ybreaks <- seq(0, round_any(max(data_bc$below, na.rm=TRUE), 1000, ceiling), 1000)
xbreaks <- seq(floor_date(min(data_bc$date_cut, na.rm=TRUE), "years"), 
               ceiling_date(max(data_bc$date_cut, na.rm=TRUE), "years"), by="1 year")
data_bc %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight (kgDM/ha)', fill='Species',
         title=paste('Mass Below Cut (below 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_point(mapping=aes(x=date_cut, y=below), colour='brown') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
  )
```

## Pasture Cuts DM%

Average cut dry matter % = `r round(mean(data_cut$dm_pc, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot DM% data
ybreaks <- seq(0, round_any(max(data_cut$dm_pc, na.rm=TRUE), 10, ceiling), 10)
xbreaks <- seq(floor_date(min(data_cut$date_cut, na.rm=TRUE), "years"), 
               ceiling_date(max(data_cut$date_cut, na.rm=TRUE), "years"), by="1 year")
data_cut %>%
  split(.$block) %>%
  map(~ggplot(.) +
  labs(x='Date', y='Dry matter %', 
       title=paste('Pregazing Dry Matter % (cut to 4cm), Block', unique(.$block))) +
  theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
  geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
  geom_point(mapping=aes(x=date_cut, y=dm_pc), colour='purple') +
  facet_grid(cultivar ~ seed_rate ) +
  scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
  scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
  )
```

## Tiller Density

Average tiller density = `r round(mean(data_till$tillers, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot tiller data 
ybreaks <- seq(0, round_any(max(data_till$tillers, na.rm=TRUE), 5000, ceiling), 5000)
xbreaks <- seq(floor_date(min(data_till$date_till), "years"), 
               ceiling_date(max(data_till$date_till), "years"), by="1 year")
data_till %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Tiller density'~(tillers~m^{2}), 
         title=paste('Tiller Density, Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
    geom_point(mapping=aes(x=date_till, y=tillers), colour='darkgreen') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, 
                       limits=c(ybreaks[1], round_any(max(data_till$tillers, na.rm=TRUE), 1000, ceiling)))
  )
```

## Botanical Composition

```{r echo=FALSE, message=FALSE, results="hide"}
# plot botanical data
ybreaks <- seq(0, 100, 20)
xbreaks <- seq(floor_date(min(data_bot2$date_bot), "years"), 
               ceiling_date(max(data_bot2$date_bot), "years"), by="1 year")
data_bot2 %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight %', fill='Species',
         title=paste('Botanical composition (cut to 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_bar(mapping=aes(x=date_bot, y=fraction, fill=species), stat='identity') +
    scale_fill_brewer(palette='Paired') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=range(xbreaks)) + # FIXME drops data?
    scale_y_continuous(breaks=ybreaks, limits=c(0, 120))
  )
```

## Ryegrass Fraction

Ryegrass fraction calculated on green mass only

```{r echo=FALSE, message=FALSE, results="hide"}
# plot ryegrass fraction
ybreaks <- seq(0, 100, 20)
xbreaks <- seq(floor_date(min(data_bot$date_bot), "years"), 
               ceiling_date(max(data_bot$date_bot), "years"), by="1 year")
data_bot %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight %', 
         title=paste('Ryegrass Fraction (above 4cm), Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_point(mapping=aes(x=date_bot, y=rgfrac), colour="darkgreen") +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=range(xbreaks)) +
    scale_y_continuous(breaks=ybreaks, limits=c(0, 120))
  )
```

## Estimate Botancial Mass

Total mass at botanical date assumed to be equal to pregraze RPM

Cut mass at botanical date assumed to be equal to cut yield

Botanical composition below cutting height estimated from Tozer data

```{r echo=FALSE, message=FALSE, results="hide"}
# plot botanical data
ybreaks <- seq(0, round_any(max(data_bm2$yield_bot, na.rm=TRUE), 1000, ceiling), 1000)
ybreaks2 <- seq(0, round_any(max(data_bm2$yield_bot, na.rm=TRUE), 100, ceiling), 100)
xbreaks <- seq(floor_date(min(data_bm2$date_bot), "years"), 
               ceiling_date(max(data_bm2$date_bot), "years"), by="1 year")
data_bm2 %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Dry weight (kgDM/ha)', fill='Species',
         title=paste('Herbage Mass Fractions, Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_point(mapping=aes(x=date_bot, y=yield_bot), colour='grey', size=2, shape=1) +
    geom_bar(mapping=aes(x=date_bot, y=species_mass, fill=species), stat='identity') +
    scale_fill_brewer(palette='Paired') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=year(xbreaks), limits=range(xbreaks)) +
    scale_y_continuous(breaks=ybreaks, limits=range(ybreaks2))
  )
```

## Soil Moisture

Average soil moisture = `r round(mean(data_sm$mean_sm, na.rm=TRUE))`

```{r echo=FALSE, message=FALSE, results="hide"}
# plot soil moisture data
ybreaks <- seq(0, round_any(max(data_sm$mean_sm, na.rm=TRUE), 20, ceiling), 20)
xbreaks <- seq(floor_date(min(data_sm$date_sm), "years"), 
               ceiling_date(max(data_sm$date_sm), "years"), by="1 year")
data_sm %>%
  split(.$block) %>%
  map(~ggplot(.) +
    labs(x='Date', y='Soil Moisture %', 
         title=paste('Soil Moisture 2017, Block', unique(.$block))) +
    theme(axis.text.x=element_text(angle=90, vjust=0.5, hjust=1)) +
    geom_vline(data=data_rpm, mapping=aes(xintercept=date_grazed), colour='grey') +
    geom_point(mapping=aes(x=date_sm, y=mean_sm), colour='blue') +
    facet_grid(cultivar ~ seed_rate ) +
    scale_x_datetime(breaks=xbreaks, labels=format(xbreaks, "%b-%Y"), limits=c(xbreaks[1], tail(xbreaks, 1))) +
    scale_y_continuous(breaks=ybreaks, limits=c(ybreaks[1], tail(ybreaks, 1)))
  )
```

